---
title: "Untitled"
author: "David Timewell"
date: "20/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,load libraries}

library(tidyverse)
library(here)
library(skimr)
library(lubridate)
library(ggplot2)
library(egg)
library(sqldf)
library(broom)
library(janitor)
library(Metrics)
library(fastDummies)
library(Amelia)
library(rsample)
library(ranger)
library(randomForest)
library(caret)
library(ROCR)
library(gbm)
library(vip)
library(corrplot)
library(ROSE)
library(broom)
 


```




```{r,import data}

# Import Data


 



data_cred <- read_csv(here("data", "AT2_credit_train.csv"))


 

```

Column 	Description
ID 	ID of each client
LIMIT_BAL 	Amount of given credit in dollars (includes individual and family/supplementary credit
SEX 	Gender (1=male, 2=female)
EDUCATION 	(1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
MARRIAGE 	Marital status (1=married, 2=single, 3=others)
AGE 	Age in years
PAY_X 	Repayment status for the past X months.
-1=paid on time; 1=payment delay for one month; 2=payment delay for two months etc
BILL_AMTX 	Bill amount for past X months
PAY_AMTX 	Payment amount for last X months
default 	Default payment next month (1=yes, 0=no). This is the Target variable



```{r}


skim(data_cred)

```



```{r}


str(data_cred)

```

# COLUMNS


## Defaut Column

Data a little imbalanced. Only 26% Y.  Consider up or down sampling in train set.

```{r}

prop.table(table(data_cred$default))


```



## SEX column

SEX 	Gender (1=male, 2=female)

```{r}

data_cred$SEX <- as.factor(data_cred$SEX) 

table(data_cred$SEX)


ggplot(data_cred,aes(x=SEX)) + geom_bar()


#NA and another column = 2113

#consider removing columns below

data_cred %>% filter(SEX %in% c(2113,NA)) 



```


## AGE column

AGE 	Age in years


```{r}

# Look at min age = 21
min(data_cred$AGE)
# look at max age = 174
max(data_cred$AGE) 




#data_cred$AGE <- as.factor(data_cred$AGE) 

#table(data_cred$AGE)


ggplot(data_cred,aes(x=AGE_BAND)) + geom_bar()


```

```{r}
#NA and another column = 2113

#consider removing columns for 100+, only 18 rows

data_cred %>% filter(AGE_BAND == "100+")


```

## Marriage

MARRIAGE 	Marital status (1=married, 2=single, 3=others)


```{r}


ggplot(data_cred,aes(x=MARRIAGE)) + geom_bar()


```


```{r}

data_cred %>% filter(MARRIAGE==0)

```

rows look sensible

opions move 0 to 3
or try impute values 



## Education



EDUCATION 	(1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)


```{r}

ggplot(data_cred,aes(x=EDUCATION)) + geom_bar()


```


```{r}



ggplot(data_cred,aes(x=EDUCATION)) + geom_bar()

```



## PAY_X


PAY_X 	Repayment status for the past X months.
-1=paid on time; 1=payment delay for one month; 2=payment delay for two months etc


```{r}

#What is -2 ?

sort(unique(data_cred$PAY_0))
#unique(data_cred$PAY_1) no PAY_1
sort(unique(data_cred$PAY_2))
sort(unique(data_cred$PAY_3))
sort(unique(data_cred$PAY_4))
sort(unique(data_cred$PAY_5))
sort(unique(data_cred$PAY_6))
```

```{r}
#There are alot of -2 rows. Impute ? or just leave them 
data_cred %>% filter(PAY_0 == -2 | PAY_2 == -2 | PAY_3 == -2 | PAY_4 == -2 | PAY_5 == -2 |  PAY_6 == -2)


```


## Limit_Bal


Amount of given credit in dollars (includes individual and family/supplementary credit




```{r}
# -99 doesn't seem right

min(data_cred$LIMIT_BAL)
max(data_cred$LIMIT_BAL)

```

```{r}

# data seems off for -99 looking at age groups. Only 19 rows maybe remove them
data_cred %>% filter(LIMIT_BAL==-99)


```

Some outliers shown in box plot

```{r}

ggplot(data_cred,aes(x=LIMIT_BAL)) + geom_boxplot()

```

## PAY_AMT_X


PAY_AMT1
```{r}
# These only have two vaues per column 0 and another value. Its roughly 20% 0 and 80% he other vaue split per column

table(data_cred$PAY_AMT1)

prop.table(table(data_cred$PAY_AMT1))



table(data_cred$PAY_AMT2)

prop.table(table(data_cred$PAY_AMT2))


table(data_cred$PAY_AMT3)

prop.table(table(data_cred$PAY_AMT3))


table(data_cred$PAY_AMT4)

prop.table(table(data_cred$PAY_AMT4))


table(data_cred$PAY_AMT5)

prop.table(table(data_cred$PAY_AMT5))




table(data_cred$PAY_AMT6)

prop.table(table(data_cred$PAY_AMT6))



```


## BILL_AMT_X

Bill amount for past X months

BILL_AMT6,4,3,1 contain negative values




```{r}


ggplot(data_cred,aes(x=BILL_AMT1)) + geom_boxplot()


```


```{r}


data_cred %>% filter(BILL_AMT1 < 0 )


```



```{r}


ggplot(data_cred,aes(x=BILL_AMT2)) + geom_boxplot()


```


```{r}


data_cred %>% filter(BILL_AMT2 < 0 )


```




```{r}


ggplot(data_cred,aes(x=BILL_AMT3)) + geom_boxplot()


```



```{r}


data_cred %>% filter(BILL_AMT3 < 0 )


```


```{r}


ggplot(data_cred,aes(x=BILL_AMT4)) + geom_boxplot()


```


```{r}


data_cred %>% filter(BILL_AMT4 < 0 )


```



```{r}


ggplot(data_cred,aes(x=BILL_AMT5)) + geom_boxplot()


```


```{r}


data_cred %>% filter(BILL_AMT5 < 0 )


```


```{r}


ggplot(data_cred,aes(x=BILL_AMT6)) + geom_boxplot()


```






```{r}


data_cred %>% filter(BILL_AMT6 < 0 )


```


## look at nearzero values


```{r Example of nearZeroVar() function}

# PAY_AMT1 - 3 

nearZeroVar(data_cred, freqCut = 95/5, uniqueCut = 10)
nearZeroVar(data_cred, freqCut = 80/20, uniqueCut = 10)
```


```{r}

str(data_cred)




```



# CHANGES



Keep original data set create a new one as the one for modelling

```{r}

model_data <- data_cred 


```




## feature changes



```{r}
#AGE
#Create age band

#Convet age to band, add as new column
labs <- c(paste(seq(0, 95, by = 5), seq(0 + 5 - 1, 100 - 1, by = 5),
                sep = "-"), paste(100, "+", sep = ""))

AGE_BAND <- cut(model_data$AGE, breaks = c(seq(0, 100, by = 5), Inf), labels = labs, right = FALSE)


model_data <- model_data %>%  mutate(AGE_BAND=as.factor(AGE_BAND))


#SEX
#remove the below columns

model_data <- model_data %>% filter(!SEX %in% c(2113,NA)) 


#MARRIAGE


# move 0 to 3

model_data$MARRIAGE[model_data$MARRIAGE %in% c(0)] <-   3


#EDUCATION


# group 0,4,5 and 6 as other



model_data$EDUCATION[model_data$EDUCATION %in% c(0,5,6)] <-   4



```


```{r}

str(model_data)

```



```{r}


model_data$SEX <- as.factor(model_data$SEX) 
model_data$EDUCATION <- as.factor(model_data$EDUCATION) 
model_data$MARRIAGE <- as.factor(model_data$MARRIAGE) 
model_data$AGE_BAND <- as.factor(model_data$AGE_BAND) 

#da_repurch <- data_repurch[ , -which(names(data_repurch) %in% c("ID"))]
model_data$default <- as.factor(model_data$default)


```

# Remove age and id 

```{r}

model_data <- model_data[ , -which(names(model_data) %in% c("ID"))]
model_data <- model_data[ , -which(names(model_data) %in% c("AGE"))]


```


# test and train splits

```{r}
set.seed(42)
sample_set <- createDataPartition(y = model_data$default, p = .70, list = FALSE)
data_train <- model_data[sample_set,]

data_test <- model_data[-sample_set,]

```


Try downsampling with cross validation

```{r}

my_control_cv_down <- trainControl(method = "cv",
                               number = 5,
                               search="grid",
                               summaryFunction = twoClassSummary,
                               classProbs = TRUE,
                               allowParallel = TRUE,
                              sampling = "down"
                               )

```


# use tune length 50 is a bit pointless as we don't have 50 variables. Caret actually picks mtry = 23

```{r}

model_train_rf_down <- train(
 default ~ .,
 data = data_train,
 metric = "ROC",
 method = "rf",
 trControl = my_control_cv_down,
 #tuneGrid = my_glmnet_grid,
 tunelength = 50
 )



```




```{r}

model_train_rf_down

```

```{r}

str(data_test)

```
# confusion matrix 

```{r}

model_rf_pred_down <- predict(model_train_rf_down, data_test)
confusionMatrix(model_rf_pred_down,  data_test$default, positive = "Y", mode = "everything")



```



# Prepare the RF submission

```{r}


valid_data <- read_csv(here("data", "AT2_credit_test.csv"))


```





# Prepare Data

```{r}
#AGE
#Create age band

#Convet age to band, add as new column
labs <- c(paste(seq(0, 95, by = 5), seq(0 + 5 - 1, 100 - 1, by = 5),
                sep = "-"), paste(100, "+", sep = ""))

AGE_BAND <- cut(valid_data$AGE, breaks = c(seq(0, 100, by = 5), Inf), labels = labs, right = FALSE)


valid_data <- valid_data %>%  mutate(AGE_BAND=as.factor(AGE_BAND))


#SEX
#remove the below columns

valid_data <- valid_data %>% filter(!SEX %in% c(2113,NA)) 


#MARRIAGE


# move 0 to 3

valid_data$MARRIAGE[valid_data$MARRIAGE %in% c(0)] <-   3


#EDUCATION


# group 0,4,5 and 6 as other



valid_data$EDUCATION[valid_data$EDUCATION %in% c(0,5,6)] <-   4


#valid_data <- valid_data[ , -which(names(valid_data) %in% c("ID"))]
valid_data <- valid_data[ , -which(names(valid_data) %in% c("AGE"))]

valid_data$SEX <- as.factor(valid_data$SEX) 
valid_data$EDUCATION <- as.factor(valid_data$EDUCATION) 
valid_data$MARRIAGE <- as.factor(valid_data$MARRIAGE) 
valid_data$AGE_BAND <- as.factor(valid_data$AGE_BAND) 



```







```{r}

model_valid_pred <- predict(model_train_rf_down, valid_data[,-1],type="prob")

```





```{r}



submission_1_rf <- tibble("ID" = valid_data$ID,'default'=model_valid_pred$Y)



write_csv(submission_1_rf,here("data", "submission_1_rf1.csv"))

```







